Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-12-25 Paper/ms.tex
minor edits

\subsection{Parameters and Priors}
To fit a planet model to RV data, we need parameters to describe
the properties of each planet. For simplicity, we describe each planet by
five parameters: the orbital period $P_i$, the semi-amplitude (in metres
per second) of the RV signal $A_i$, the phase of the signal
$\phi_i$, the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$.

@@

The unknown parameters are:
\begin{eqnarray}
\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0, \sigma_{\rm extra}, \nu\right\}
\end{eqnarray}
where $N$ is the number of planets,
$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are the
hyperparameters used to define the prior for the properties of the planets,
and $\psi_i = \{P_i, A_i, \phi_i, e_i, \theta_i\}$
are the properties of planet $i$.
The parameter $m_0$ describes a DC offset in the data, and
$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution.

@@

All of the model assumptions are specified in detail in
Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
prior knowledge, rather than the judgement of an informed expert on extrasolar
planets. Uniform distributions were used for parameters such as phases, where
time-translation symmetry seems plausible.
Truncated Cauchy distributions were used when there is a
preferred value, but since these have very heavy tails, the assumption is
quite fail safe relative to other possible assignments such as normal distributions.
For example, the prior for $\mu_P$, the typical orbital period,
is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
as $\approx e^{21}$ years, a very wide range. Truncation intervals are
specified with the $T(a,b)$ notation in Table~\ref{tab:priors}.

An apparently strange choice is the conditional prior for the orbital
periods, which is a biexponential distribution. A more conventional choice
would have been a normal distribution. However, the \citet{rjobject} software
needs to know the corresponding cumulative distribution and its inverse, which
are not available in closed form for the normal distribution.

The prior for the velocity semi-amplitudes $\{A_i\}$ is important and will affect the
inferred number of planets $N$. For example, if the prior allows for very
low amplitude planets to exist, the data will not rule this out, and the posterior
for $N$ will include the possibility of large numbers of planets. The prior
will influence how many of these low amplitude planets will exist: if we
believe there are many, and the data are uninformative about low amplitude
planets, then the posterior distribution for $N$ will also indicate that there
may be many low amplitude planets. However, their other properties, such as
their orbital periods, will not be well determined.

>>>>

\subsection{Parameters and Priors}
To fit a planet model to RV data, we need parameters to describe
the properties of each planet. For simplicity, we describe each planet by
five parameters: the orbital period $P_i$, the semi-amplitude (in metres
per second) of the RV signal $A_i$, the phase of the signal
$\phi_i$ (defined such that $\psi=0$ gives an RV signal whose maximum is at
$t=0$), the eccentricity $e_i$, and the ``viewing angle'' $\theta_i$
(also known as the longitude of the line of sight).

@@

The unknown parameters are:
\begin{eqnarray}
\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0, \sigma_{\rm extra}, \nu\right\}
\end{eqnarray}
where $N$ is the number of planets,
$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are the
hyperparameters used to define the prior for the properties of the planets,
and $\psi_i = \{P_i, A_i, e_i, \phi_i, \theta_i\}$
are the properties of planet $i$.
The parameter $m_0$ describes a DC offset in the data, and
$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution
which are discussed further below.

@@

All of the model assumptions are specified in detail in
Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
prior knowledge, rather than the judgement of an informed expert on extrasolar
planets. Uniform distributions were used for parameters such as phases, where
time-translation symmetry seems plausible.
Truncated Cauchy distributions were used when there is a
preferred value, but since these have very heavy tails, the assumption is
quite fail safe relative to other possible assignments such as normal distributions.
For example, the prior for $\mu_P$, the typical orbital period,
is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
as $\approx e^{21}$ years, a very generous range. Truncation intervals are
specified with the $T(a,b)$ notation in Table~\ref{tab:priors}.

An apparently strange choice is the conditional prior for the orbital
periods, which is a biexponential distribution. A more conventional choice
would have been a normal distribution. However, the \citet{rjobject} software
needs to know the corresponding cumulative distribution and its inverse, which
are not available in closed form for the normal distribution.
The prior for the velocity semi-amplitudes $\{A_i\}$ is important and will affect the
inferred number of planets $N$. For example, if the prior allows for very
low amplitude planets to exist, the data will not rule this out, and the posterior
for $N$ will include the possibility of large numbers of planets. The prior
will influence how many of these low amplitude planets will exist: if we
believe there are many, and the data are uninformative about low amplitude
planets, then the posterior distribution for $N$ will also indicate that there
may be many low amplitude planets. However, their other properties, such as
their orbital periods, will not be well determined.
